---
title: "Map of Recent Earthquakes Around Italy"
author: "Alessandro Vullo"
date: "16 October 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview 

The aim in this report is to create and display an interactive map of recent earthquakes around Italy and the surrounding seas. Data is taken from the [Italian National Institute of Geophysics and Volcanology](http://www.ingv.it/en/) which provides detailed information about earthquakes available to download.

## Data Processing

### Getting the Raw Data

Data available at the [INGV page](http://info.terremoti.ingv.it/) is taken using the [INGV webservices](http://webservices.ingv.it) and downloaded as a text file to a local file.

```{r}
if(!file.exists('earthquakes.txt')) {
  download.file('http://webservices.ingv.it/fdsnws/event/1/query?starttime=2017-10-09T00%3A00%3A00&endtime=2017-10-16T23%3A59%3A59&minmag=2&maxmag=10&mindepth=-10&maxdepth=1000&minlat=-90&maxlat=90&minlon=-180&maxlon=180&minversion=100&orderby=time-asc&format=text&limit=10000', destfile = 'earthquakes.txt', method = "curl")
}
```

### Reading the Data

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(lubridate)
library(leaflet)
```

The downloaded file is a simple text with header and columns delimited by '|'.

```{r loaddata}
df <- read.csv('earthquakes.txt', header=TRUE, sep="|")
str(df)
```

### Filtering Data

We add a column to provide the earthquake location, time and magnitude as popup information when selected by the user on the map:

```{r}
df <- df %>%
      mutate(popup = paste(sep = "<br>", 
                           EventLocationName, 
                           Time, 
                           paste(sep = " ", MagType, Magnitude)))
```

## Interactive Earthquakes Map

Finally, the map is rendered where each earthquake is marked by a circle with radius proportional to its magnitude squared. The earthquakes are clustered on the map:

```{r map}
leaflet(df) %>% 
  addTiles() %>%
  setView(13,41.5,zoom=6) %>%
  addCircleMarkers(lat = ~Latitude, lng = ~Longitude, 
                   radius = ~Magnitude^2, 
                   clusterOptions = markerClusterOptions(), 
                   popup = ~as.character(popup))
```
